---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
---



```{r}

setwd("C:\\Users\\jyoti\\OneDrive\\Documents\\PGPBABI\\Data Mining")
creditdata = read.csv('g1.csv')
str(creditdata)
dim(creditdata)
```

#Convert credit to categorical variable
```{r}
creditdata$CREDIT = as.factor(creditdata$CREDIT)
names(creditdata)

```

#Find Baseline Y class %
```{r}
prop.table(table(creditdata$CREDIT))
```


### Exploratory data analysis - boxplot
```{r}
library(ggplot2)

ggplot(data=creditdata, aes(x=CREDIT, y=DURATION, fill=CREDIT))+geom_boxplot()

ggplot(data=creditdata, aes(x=CREDIT, y=AMOUNT, fill=CREDIT))+geom_boxplot()

ggplot(data=creditdata, aes(x=CREDIT, y=AGE, fill=CREDIT))+geom_boxplot()



```


#Stacked bar plot for categorical variables
```{r}
ggplot(creditdata, aes(fill=CREDIT, x=CHECK_A, y=as.numeric(CHECK_A))) +
  geom_bar( stat="identity", position="fill")

ggplot(creditdata, aes(fill=CREDIT, x=CHECK_A, y=as.numeric(CHECK_A))) +
  geom_bar( stat="identity")


ggplot(creditdata, aes(fill=CREDIT, x=C_HIST, y=as.numeric(C_HIST))) +
  geom_bar( stat="identity", position="fill")



ggplot(creditdata, aes(fill=CREDIT, x=PURPOSE, y=as.numeric(PURPOSE))) +
  geom_bar( stat="identity", position="fill")



ggplot(creditdata, aes(fill=CREDIT, x=SAVE_A, y=as.numeric(SAVE_A))) +
  geom_bar( stat="identity", position="fill")


```



```{r}
## split into training and test sets
library(caTools)

set.seed(123)
split = sample.split(creditdata$CREDIT, SplitRatio = 0.80)
traindata = subset(creditdata, split == TRUE)
testdata = subset(creditdata, split == FALSE)

## Check if distribution of partition data is correct Testing dataset
prop.table(table(creditdata$CREDIT))
prop.table(table(creditdata$CREDIT))
```

Build a fully grown tree
```{r}
library(rpart)
library(rpart.plot)
set.seed(420)

tree_full = rpart(formula = CREDIT~., data = traindata, cp=-1, minsplit=2, minbucket=1)
rpart.plot(tree_full, cex=0.4)

boxcols <- c("palegreen3", "orange")[tree_full$frame$yval]

par(xpd=TRUE)
prp(tree_full, faclen = 0, cex = 0.3, extra = 1, box.col = boxcols)
```


```{r}
print(tree_full)
```


#predict on train and test set
```{r}
## Predict using the CART model
traindata$predict.class=predict(tree_full,traindata,type="class")
traindata$predict.score=predict(tree_full,traindata)

## Creating the confusion matrix
tabtrain=with(traindata,table(CREDIT,predict.class))
tabtrain

TN_train = tabtrain[1,1]
TP_train = tabtrain[2,2]
FN_train = tabtrain[2,1]
FP_train = tabtrain[1,2]

train_acc = (TN_train+TP_train)/(TN_train+TP_train+FN_train+FP_train)
train_acc

train_sens = TP_train/(TP_train+FN_train)
train_sens


train_spec = TN_train/(TN_train+FP_train)
train_spec
```




```{r}
## Predict using the CART model
testdata$predict.class=predict(tree_full,testdata,type="class")
testdata$predict.score=predict(tree_full,testdata)

## Creating the confusion matrix
tabtest=with(testdata,table(CREDIT,predict.class))
tabtest

TN_test = tabtest[1,1]
TP_test = tabtest[2,2]
FN_test = tabtest[2,1]
FP_test = tabtest[1,2]

test_acc = (TN_test+TP_test)/(TN_test+TP_test+FN_test+FP_test)
test_acc

test_sens = TP_test/(TP_test+FN_test)
test_sens


test_spec = TN_test/(TN_test+FP_test)
test_spec
```


```{r}
df_results_train = data.frame(train_acc, train_sens, train_spec)
names(df_results_train) = c("ACC", "SENS", "SPEC")
df_results_test = data.frame(test_acc, test_sens, test_spec)
names(df_results_test) = c("ACC", "SENS", "SPEC")

?rbind
df_fin =rbind(df_results_train, df_results_test)
row.names(df_fin) = c('tree_full_train', 'tree_full_test')
df_fin

```


#remove predicted score and class before running other models
```{r}
traindata$predict.class = NULL
traindata$predict.score = NULL
testdata$predict.class = NULL
testdata$predict.score = NULL
```




#Pruning using minbucket and minsplit
```{r}

set.seed(123)
?rpart

tree_manual_prune=rpart(formula = CREDIT ~ ., data = traindata, method="class",control = rpart.control(minsplit = 50,  minbucket = 15))



rpart.plot(tree_manual_prune, cex=0.8)
```


```{r}
## Predict using the CART model
traindata$predict.class=predict(tree_manual_prune,traindata,type="class")
traindata$predict.score=predict(tree_manual_prune,traindata)

## Creating the confusion matrix
tabtrain=with(traindata,table(CREDIT,predict.class))
tabtrain

TN_train = tabtrain[1,1]
TP_train = tabtrain[2,2]
FN_train = tabtrain[2,1]
FP_train = tabtrain[1,2]

train_acc = (TN_train+TP_train)/(TN_train+TP_train+FN_train+FP_train)
train_acc

train_sens = TP_train/(TP_train+FN_train)
train_sens


train_spec = TN_train/(TN_train+FP_train)
train_spec
```


```{r}
## Predict using the CART model
testdata$predict.class=predict(tree_manual_prune,testdata,type="class")
testdata$predict.score=predict(tree_manual_prune,testdata)

## Creating the confusion matrix
tabtest=with(testdata,table(CREDIT,predict.class))
tabtest

TN_test = tabtest[1,1]
TP_test = tabtest[2,2]
FN_test = tabtest[2,1]
FP_test = tabtest[1,2]

test_acc = (TN_test+TP_test)/(TN_test+TP_test+FN_test+FP_test)
test_acc

test_sens = TP_test/(TP_test+FN_test)
test_sens


test_spec = TN_test/(TN_test+FP_test)
test_spec
```


```{r}
df_results_train = data.frame(train_acc, train_sens, train_spec)
names(df_results_train) = c("ACC", "SENS", "SPEC")
df_results_test = data.frame(test_acc, test_sens, test_spec)
names(df_results_test) = c("ACC", "SENS", "SPEC")

?rbind
df_fin =rbind(df_fin, df_results_train, df_results_test)
row.names(df_fin) = c('tree_full_train', 'tree_full_test', 'tree_train', 'tree_test')
df_fin
```


```{r}
#remove predicted score and class before running other models
traindata$predict.class = NULL
traindata$predict.score = NULL
testdata$predict.class = NULL
testdata$predict.score = NULL
```


#Prune using cp
```{r}

printcp(tree_full)
plotcp(tree_full)

bestcp=tree_full$cptable[which.min(tree_full$cptable[,"xerror"]),"CP"]
bestcp


ptree=prune(tree_full,cp=bestcp)
print(ptree)
rpart.plot(ptree, cex = 0.8)
```


```{r}
## Predict using the CART model
traindata$predict.class=predict(ptree,traindata,type="class")
traindata$predict.score=predict(ptree,traindata)

## Creating the confusion matrix
tabtrain=with(traindata,table(CREDIT,predict.class))
tabtrain

TN_train = tabtrain[1,1]
TP_train = tabtrain[2,2]
FN_train = tabtrain[2,1]
FP_train = tabtrain[1,2]

train_acc = (TN_train+TP_train)/(TN_train+TP_train+FN_train+FP_train)
train_acc

train_sens = TP_train/(TP_train+FN_train)
train_sens


train_spec = TN_train/(TN_train+FP_train)
train_spec

## Predict using the CART model
testdata$predict.class=predict(ptree,testdata,type="class")
testdata$predict.score=predict(ptree,testdata)

## Creating the confusion matrix
tabtest=with(testdata,table(CREDIT,predict.class))
tabtest

TN_test = tabtest[1,1]
TP_test = tabtest[2,2]
FN_test = tabtest[2,1]
FP_test = tabtest[1,2]

test_acc = (TN_test+TP_test)/(TN_test+TP_test+FN_test+FP_test)
test_acc

test_sens = TP_test/(TP_test+FN_test)
test_sens


test_spec = TN_test/(TN_test+FP_test)
test_spec
```


```{r}
df_results_train = data.frame(train_acc, train_sens, train_spec)
names(df_results_train) = c("ACC", "SENS", "SPEC")
df_results_test = data.frame(test_acc, test_sens, test_spec)
names(df_results_test) = c("ACC", "SENS", "SPEC")

?rbind
df_fin =rbind(df_fin, df_results_train, df_results_test)
row.names(df_fin) = c('tree_full_train', 'tree_full_test', 'tree_train', 'tree_test', 'ptree_train', 'ptree_test')
round(df_fin,2)
```
#Best result from tree with manual pruning



```{r}
boxcols <- c("palegreen3", "orange")[tree_full$frame$yval]

par(xpd=TRUE)
prp(ptree, faclen = 0, cex = 0.6, extra = 1, box.col = boxcols)



```



```{r}
#Variable importance
library(caret)

ptree$variable.importance

df_cart=data.frame(round(ptree$variable.importance,2))
df_cart
write.csv(df_cart,"varimpcart.csv")
```


```{r}
traindata$predict.class = NULL
traindata$predict.score = NULL
testdata$predict.class = NULL
testdata$predict.score = NULL
```



#Calculate AUC and KS with best model
```{r}
## Predict using the CART model
traindata$predict.class=predict(tree_manual_prune,traindata,type="class")
traindata$predict.score=predict(tree_manual_prune,traindata)

View(traindata)

```


```{r}
library(pROC)
roc_obj = roc(traindata$CREDIT, traindata$predict.score[,2])


plot(roc_obj, print.auc = T)
```


