---
title: "Hierarchical Clustering"
output: html_notebook
---

Use Eng_colg_Data. Here we have data of 26 colleges collected after Survey Research using questionnaires.
As it is evident, the initial data has been analyzed and aggregated to an average rating for each college.
The ratings are in the standardized scale of 1 to 5 where 5 has a higher weightage than 1.Use your best judgment and build the Clustering Model on this data.
 
The Report should be properly formatted and should include the following :-
What do you make out of the given data?
What will be your approach to solve this problem? (you are advised to give supporting arguments for the methodology you are going to use)
What are the findings? ( visualize your clusters, screen-shot of your output)
Interpretation of your findings 
Appendix: Paste the code at the end

```{r}
setwd("C:\\Users\\jyoti\\OneDrive\\Documents\\PGPBABI\\Data Mining\\clustering_LMS")
engdata <- read.csv("Engg_College_Data.csv", header=TRUE)
engdata$SR_NO = NULL
View(engdata)
```

##Calculate euclidean distance
```{r}
d_euc <- dist(x=engdata[,3:6], method = "euclidean") 
print(d_euc, digits = 3)
```


##Use hierarchical clustering
```{r}
clus1 <- hclust(d_euc, method = "average")
plot(clus1, labels = as.character(engdata[,2]))
rect.hclust(clus1, k=5, border="red")

```

## profiling the clusters
```{r}

?cutree

engdata$Clusters <- cutree(clus1, k=5)

View(engdata)

aggr = aggregate(engdata[,-c(1,7)],list(engdata$Clusters),mean)
aggr

clus.profile <- data.frame( Cluster=aggr[,1],
                            Freq=as.vector(table(engdata$Clusters)),
                            aggr[,-1])

View(clus.profile)
```



